---
title: "PATRIC_homology_3cyano"
output: html_document
date: "2023-06-21"
---
Load in data
```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(data.table)

loci <- read_csv("locus_03_04_2024.csv")


homology <- read.csv("Homology_7942.csv")


expression_6803 <- read_csv("6803logTPM_counts_051623.csv")
expression_7002 <- read_csv("7002log_tpm_counts.csv")
expression_7942 <- read_csv("7942log_TPM_counts_zach.csv")
```

Map the values from the old locus tags to the new locus tags. Some of the tags have no mapping, ex. sml0003 in 6803. We identify the invalid loci. 
```{r, message=FALSE}
new <- loci$locus_tag
old <- loci$old_locus_tag

#mapping old to new locus tags
pcc6803 <- plyr::mapvalues(homology$comp_genome_2_locus_tag, old, new)
pcc7002 <- plyr::mapvalues(homology$comp_genome_1_locus_tag, old, new)
pcc7942 <- plyr::mapvalues(homology$ref_genome_locus_tag, old, new)

#find the invalid loci which do not have standard new naming found in expression data
invalid_6803 <- which(!grepl("SGL_RS", pcc6803))
invalid_7002 <- which(!grepl("SYNPCC7002_RS", pcc7002))
invalid_7942 <- which(!grepl("SYNPCC7942_RS", pcc7942))

#vector of the names of loci removed from each group 
removed_6803 <- pcc6803[invalid_6803]
removed_7002 <- pcc7002[invalid_7002]
removed_7942 <- pcc7942[invalid_7942]

#removing invalid loci from the homologous genes. Doesn't really do anything 
#pcc6803 <- pcc6803[-invalid_6803]
#pcc7002 <- pcc7002[-invalid_7002]
#pcc7942 <- pcc7942[-invalid_7942]
```

Tie the genes into triplets based on homology
```{r}
compare <- lapply(1:length(homology$ref_genome_locus_tag), function(x) {c(homology$comp_genome_2_locus_tag[x], homology$comp_genome_1_locus_tag[x], homology$ref_genome_locus_tag[x])})
```

We use the empty strings, invalid loci, and loci not found in expression data to help prune the triplets (compare) down to the usable set of loci. 
```{r}
helper <- function(triplet){
  n1 <- ifelse(triplet[1]=="", "not available", new[which(old == triplet[1])])
  n2 <- ifelse(triplet[2]=="", "not available", new[which(old == triplet[2])])
  n3 <- ifelse(triplet[3]=="", "not available", new[which(old == triplet[3])])
 
  #Need to check that none are empty strings, none are removed, and that none are not found in the expression data. Some genes are in homology but not in expression. 
  if(any(triplet == "") | triplet[1] %in% removed_6803 | triplet[2] %in% removed_7002 | triplet[3] %in% removed_7942 | !n1 %in% expression_6803$...1 | !n2 %in% expression_7002$...1 | !n3 %in% expression_7942$...1){ 
    return(NA)
  }
  return(c(n1, n2, n3))
}

triples <- lapply(compare, helper)
triples <- triples[!is.na(triples)]
present_genes <- unlist(triples)
fin1 <- filter(expression_6803, expression_6803$...1 %in% present_genes)
fin2 <- filter(expression_7002, expression_7002$...1 %in% present_genes)
fin3 <- filter(expression_7942, expression_7942$...1 %in% present_genes)

#Because we do PATRIC on 6803, all genes from sp6803 are going to be unique. In the end there are 1383 triplets of homology and 1383 unique sp6803 genes. However 7002 and 7942 have less than 1383 unique genes in homology because of potential genes that are homologous to multiple sp6803 genes.  
sum(duplicated(present_genes)) #111, which is exactly (fin1-fin2) + (fin1-fin3) with ref 6803. 73 with ref 7942. 0 with duplicates removed from data set.

#which(duplicated(present_genes))
#present_genes[334]
#SGL_RS04780
```

Put the dataframe together to export
```{r}
#Turn the rownames into the locus tags so we can reference the row by name. Drop the locus tags for fin1 (6803) and fin2 (7002) so that we can add the expression data horizontally
fin1 <- fin1 %>% column_to_rownames(var="...1")
fin2 <- fin2 %>% column_to_rownames(var="...1")
fin3$locus = fin3$...1
fin3 <- fin3 %>% column_to_rownames(var="locus")

#Calls each gene in the triple by row merges them as a long vector  
buildTable <- function(triplet){
   return(c(as.vector(fin3[triplet[3],]), as.vector(fin2[triplet[2],]), as.vector(fin1[triplet[1],]), triplet[2], triplet[1]))
}

#It sorts the 6803 and 7002 data columns alphabetically but it does not sort 7942. V1 and V2 are the names of the 7002 and 6803 locus tag of homologous genes.  
full_expression_data_homology <- rbindlist(lapply(triples, buildTable))

write_csv(full_expression_data_homology, "genie3_unduplicated_03_20_2024.csv")

```

Ignore the rest please.

```{r}
old_tag_data <- read_csv("C:/Users/liro650/OneDrive - PNNL/Documents/PATRIC/genie3_unduplicated.csv", col_select = c(1))

new_tag_data <- read_csv("C:/Users/liro650/OneDrive - PNNL/Documents/PATRIC/genie3_unduplicated_new_locus.csv", col_select = c(1))

new <- which(!new_tag_data$...1 %in% old_tag_data$...1)
new_tag_data$...1[new]

lost <- which(!old_tag_data$...1 %in% new_tag_data$...1)
old_tag_data$...1[lost]

```





graph the network
```{r}
library(tidyverse)
library(igraph)
library(Matrix)
library(RCy3)

linkedList <- read_csv("C:/Users/liro650/OneDrive - PNNL/Documents/genie3/GenieLinkList_PNNL_Homology_7942_ds6.csv")

g <- graph.data.frame(linkedList, directed = F)
adj <- get.adjacency(g, sparse = F, attr = "weight",type="both")



```


Miscellaneous, don't run
```{r, eval = FALSE}
preserve <- which(pcc6803 %in% expression_6803$...1)
expression_6803 <- filter(expression_6803, row_number() %in% preserve)

preserve <- which(pcc7002 %in% expression_7002$...1)
expression_7002 <- filter(expression_7002, row_number() %in% preserve)

preserve <- which(pcc7942 %in% expression_7942$...1)
expression_7942 <- filter(expression_7942, row_number() %in% preserve)
```


